### 2020-06-17 text/article_vedrovice.Rmd
spatial analysis

## Spatial randomness

@@

By counting co-occurences of analyzed variables on simulated data matrices and comparing
the distributions of the two studied statistics with the values calculated for
the original dataset we can establish whether the overall pattern of artefacts
distribution in the graves is non-random and further identify artefacts that deviate
from the random pattern.
The value of $S$ statistic is deemed significantly large at the 0.05 alpha level with
$S \approx 14 $ and $ p \approx $ (Fig. \@ref(fig:f-sstat)).
This suggests that there is enough evidence that the overall distribution of
artefacts in the graves is not random.

```{r t-vstat}
vstat <- read_csv(here("data/temp", "v_statistics.csv")) %>%
    select(full, v, p) %>%
    mutate(v = round(v, 2), p = round(p, 2)) %>%
    rename("variable" = full, "v statistic" = v, "p value" = p)

knitr::kable(
  vstat,
  caption = 'Fuuuuu'
)
```

>>>>

## Spatial organization

Spatial arrangement of prehistoric burial grounds is often examined to find
any internal structures, arrangements in lines, possible grouping and so forth.
We approach the task as a point pattern analysis problem given that
a burial (event) is simplified as a point location in geographic space.
The excavated polygon is used as a study area (a window).
Using a window closely approximating locations of the burials was also considered,
e. g. buffer zone or convex/concave hull but simplified excavation polygon is
deemed the most useful because it preserves the *natural* borders and empty areas
in the burial ground.

Overall structure of the burial ground is examined using both density based approach,
and distance based methods [@osullivan2010, 121-154].
The density of burials is examined using quadrat test and the $G$ function while
the distance between neighbours usnig Ripley's $K$ function and its derivative, the
$L$ function [@ripley1976].
Given statistics for independent random process are estimated using Monte Carlo simulation
with 99 iterations and the most extreme results for each of the statistics
are used to construct envelopes encompassing the expected ranges

```{r spat-gl, fig.cap = "G and L functions"}
knitr::include_graphics(here("plots", "pointprocess_gl.pdf"))
```

@@

By counting co-occurences of analyzed variables on simulated data matrices and comparing
the distributions of the two studied statistics with the values calculated for
the original dataset we can establish whether the overall pattern of artefacts
distribution in the graves is non-random and further identify artefacts that deviate
from the random pattern.
The value of $S$ statistic is deemed significantly large at the 0.05 alpha level with
$S \approx 14$ and $p \approx xy$ (Fig. \@ref(fig:f-sstat)).
This suggests that there is enough evidence that the overall distribution of
artefacts in the graves is not random.

<!-- ```{r t-vstat} -->
<!-- vstat <- read_csv(here("data/temp", "v_statistics.csv")) %>% -->
<!--     select(full, v, p) %>% -->
<!--     mutate(v = round(v, 2), p = round(p, 2)) %>% -->
<!--     rename("variable" = full, "v statistic" = v, "p value" = p) -->

<!-- knitr::kable( -->
<!--   vstat, -->
<!--   caption = 'Fuuuuu' -->
<!-- ) -->
<!-- ``` -->
